---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Hello MLHO!

<!-- badges: start -->
<!-- badges: end -->

MLHO (pronounced as melo) is a thinkin' Machine Learning framework that implements iterative sequential representation mining, and feature and model selection to predict health outcomes.


## Installation

You can install the released version of mlho from [Github](https://https://github.com/hestiri/mlho) with:

``` r
devtools::install_github("hestiri/mlho")
```

## Data model

To implement MLHO you'll need 2 tables, which can be extracted from any clinical CMD. The current examples are based on the i2b2 star schema.

1-  a table with outcome labels (called labeldt) and patient numbers
```{r,echo=FALSE,warning=FALSE}
knitr::kable(
  rbind(c("patient_num","label"),
                   c("character","factor"))
)
```



2- a patient clinical data table (called dbmart) with 3 columns. Concepts are used as features by MLHO.

```{r,echo=FALSE,warning=FALSE}
knitr::kable(
  rbind(c("patient_num","start_date","oncept_cd"),
                   c("character","date","character"))
)
```

3- a demographic table is optional.



