---
title: "MLHO Vanilla"
author: Hossein Estiri
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MLHO Vanilla}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mlho)
```



## MLHO Vanilla implementation

Here we go through the vanilla implementation of mlho using the synthetic data provided in the package.

The synthetic data `syntheticmass` downloaded and prepared according to mlho input data model from [SyntheticMass](https://synthea.mitre.org/downloads), generated by [SyntheaTM](https://synthetichealth.github.io/synthea/), an open-source patient population simulation made available by [The MITRE Corporation](https://health.mitre.org/).

```{r}
library(mlho)
data("syntheticmass")
```
here's how `dbmart` table looks (there's an extra column that is the translation of the `phenx` concepts):
```{r}
head(dbmart)
```
our demographic table contains: `r colnames(dems)`, which we store under `dems.names`:
```{r}
dems.names <- colnames(dems)
```


### MSMR lite
from here, we will split the data into a train and a test set and apply `MSMR.lite` to the training data
