---
title: "MLHO Vanilla"
author: Hossein Estiri
date: Nov. 5th 2021
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MLHO Vanilla}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# make sure the package is installed --> devtools::install_github("hestiri/mlho")
library(mlho)
```

Install and load the required packages:
```{r}
if(!require(pacman)) install.packages("pacman")

pacman::p_load(data.table, devtools, backports, Hmisc, tidyr,dplyr,ggplot2,plyr,scales,readr,
               httr, DT, lubridate, tidyverse,reshape2,foreach,doParallel,caret,gbm,lubridate,praznik)
```

## MLHO Vanilla implementation

Here we go through the vanilla implementation of MLHO using the synthetic data provided in the package.

The synthetic data `syntheticmass` downloaded and prepared according to MLHO input data model from [SyntheticMass](https://synthea.mitre.org/downloads), generated by [SyntheaTM](https://synthetichealth.github.io/synthea/), an open-source patient population simulation made available by [The MITRE Corporation](https://health.mitre.org/).

```{r}
data("syntheticmass")
```
here's how `dbmart` table looks (there's an extra column that is the translation of the `phenx` concepts):
```{r}
head(dbmart)
```
our demographic table contains: `r colnames(dems)`.



##split data into train-test
```{r}
uniqpats <- c(as.character(unique(dbmart$patient_num)))
#using a 70-30 ratio
test_ind <- sample(uniqpats,
                   round(.3*length(uniqpats)))

test_labels <- subset(labeldt,labeldt$patient_num %in% c(test_ind))
print("test set lables:")
table(test_labels$label)
train_labels <- subset(labeldt,!(labeldt$patient_num %in% c(test_ind)))
print("train set lables:")
table(train_labels$label)
# train and test sets 
dat.train  <- subset(dbmart,!(dbmart$patient_num %in% c(test_ind)))
dat.test <- subset(dbmart,dbmart$patient_num %in% c(test_ind))
```

now dimensionality reduction on training set
## MSMR lite
from here, we will split the data into a train and a test set and apply `MSMR.lite` to the training data
```{r}
data.table::setDT(dat.train)
dat.train[,row := .I]
dat.train$value.var <- 1
uniqpats.train <- c(as.character(unique(dat.train$patient_num)))

##here is the application of MSMR.lite
dat.train <- MSMSR.lite(MLHO.dat=dat.train,
                        patients = uniqpats.train,
                        sparsity=0.005, 
                        labels = labeldt,
                        topn=200)
```
Notice that we are removing concepts that had prevalence less than 0.5% and then only taking the top 200 after the JMI rankings. See help (`?mlho::MSMSR.lite`) for `MSMR.lite` parameters. 

Now we have the training data with the top 200 features, to which we can add demographic features. Now on to prepping the test set:
```{r}
dat.test <- subset(dat.test,dat.test$phenx %in% colnames(dat.train))
setDT(dat.test)
dat.test[,row := .I]
dat.test$value.var <- 1
uniqpats.test <- c(as.character(unique(dat.test$patient_num)))

dat.test <- MSMSR.lite(MLHO.dat=dat.test,patients = uniqpats.test,sparsity=NA,jmi = FALSE,labels = labeldt)
```
Here notice that we only used `MSMR.lite` to generate the wide table that matches to the `dat.train` table.


##Modeling
we will use the `mlearn` function to do the modeling, which includes training the model and testing it on the test set.
```{r, warning=FALSE}
model.test <- mlearn(dat.train,
                   dat.test,
                   dems=dems,
                   save.model=FALSE,
                   classifier="gbm",
                   note="mlho_terst_run",
                   cv="cv",
                   nfold=5,
                   aoi="prediabetes",
                   multicore=FALSE)

```
try `?mlho::mlearn` to learn more about `mlearn` parameters. 


you can select a variety of models. For binray classification tasks, chose from:ORFlog, RRF,gbm,bayesglm,regLogistic,xgbDART,regLogistic, glmnet, bayesglm, nb, svmRadialWeights,avNNet,and ordinalRF. see the list and description of all available [models](https://rdrr.io/cran/caret/man/models.html)


## `mlearn` outputs
outputs of `mlearn` include:
  
* model performance metrics: `model.test$ROC`
  + in model.test AUROC is `r model.test$ROC$roc`
* final important features, which can be pulled and save from `model.test$features`
* absolute errors for each patients using `model.test$AE`
* missing important features `model.test$missing.features`


For instance, only the following `r nrow(model.test$features)` features were used in phenotyping the `prediabetes`:

```{r}
model.test$features
```


*we recommend iterating the training and testing and storing the coefficients in a directory.*


## Visualizing the outputs
Here we create a plot of the feature importance scores for each of the top (here we have `r nrow(model.test$coefficients)`) predictors identified by MLHO.

To do so, let's map the concept codes to their "English" translation. *That's why we kept that 4th column called `description` in `dbmart`.*
```{r}
dbmart.concepts <- dbmart[!duplicated(paste0(dbmart$phenx)), c("phenx","DESCRIPTION")]
mlho.features <- data.frame(merge(model.test$features,dbmart.concepts,by.x="features",by.y = "phenx"))
datatable(dplyr::select(mlho.features,features,DESCRIPTION,`Feature importance`=Overall), options = list(pageLength = 5), filter = 'bottom')
```

now visualizing feature importance
```{r,fig.align='center',fig.width=7,fig.height=7}
(plot<- ggplot(mlho.features) +
    geom_segment(
      aes(y = 0,
          x = reorder(DESCRIPTION,Overall),
          yend = Overall,
          xend = DESCRIPTION),
      size=0.5,alpha=0.5) +
    geom_point(
      aes(x=reorder(DESCRIPTION,Overall),y=Overall),
      alpha=0.5,size=2,color="red") +
    theme_minimal()+
   coord_flip()+
    labs(y="Feature importance",x=""))
```




## Notes for more complex implementation
As we recommended above, it's best to at least iterate the training process 5 times, in which case you'll have 5 files containing the important features from MLHO. We do this regularly. If you do so, you can compute the mean/median values for each feature.


Now if you want to get the regression coefficients and compute Odds Ratios (ORs), you can then select the top feature that were selected most frequently in MLHO iterations to run a regression model (pass `glm` to `mlearn`) and extract regression coefficients to then transform to ORs. 






