---
title: "modeling with 4CE severity codes"
author: Hossein Estiri
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{modeling with 4CE severity codes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mlho)
```
## Step 0: Data model (if you have 2.1 data extracts, skip to step 1)

To implement this analysis, you'll need 2 tables, which can be extracted from any clinical CMD. The current examples are based on the i2b2 star schema.

1-  a demographic table with outcome labels (called dems), demographic columns of interest, and patient numbers. 
The assumption is that all these patients are hospitalized for COVID-19.
```{r,echo=FALSE,warning=FALSE}
knitr::kable(
  rbind(c("","patient_num",".. demographic columns ..","label"),
                   c("TYPE","character","character","factor"))
)
```

*demographic variables need to be in binary format.*


2- a patient clinical data table (called dbmart) with 3 columns. 
```{r,echo=FALSE,warning=FALSE}
knitr::kable(
  rbind(c("patient_num","start_date","oncept_cd"),
                   c("character","date","character"))
)
```

Below is an example SQL code pulled through RODBC
```{r} 
# dbmart <- sqlQuery(odbc,paste0("select f.patient_num, f.encounter_num, f.start_date, c.concept_cd from COVID19_Mart.RPDR.OBSERVATION_FACT_RPDR f 
#                                   inner join COVID19_Mart.RPDR.VISIT_DIMENSION v on f.patient_num = v.patient_num and f.encounter_num = v.encounter_num
#                                   inner join COVID19_Mart.RPDR.CONCEPT_DIMENSION c on f.concept_cd=c.concept_cd
#                                   where f.patient_num in ('",paste(dems$patient_num,collapse = "','"),"')"))
```

### data prep and modeling.\
step 1- load the dems and datamart tables\
step 2- here we limit the dbmart concepts to 4CE severity codes. You can use a table with severity codes to subset dbmart.\
step 3- now that your data is ready, get a list of the patient numbers in your tables to use in the datebuff function.\
```{r}
#uncomment the following line to execute
#dbmart <- datebuff(data=dbmart,demographics=dems,eve="hospitallization_date",use="after",patients=uniqpats)
```

## Step 1: Load and prepare Phase 2.1 data (if you did Step 0, skip to step 2)
```{r}
#uncomment the following line to execute
#PatientMapping <- read.csv(file.choose()) ## this will open a window to select the .csv file
#PatientObservations <- read.csv(file.choose())
#PatientClinicalCourse <- read.csv(file.choose())
#PatientSummary <- read.csv(file.choose())

#severcode <- read.csv(file.choose()) ##list of severe codes
```
 
###extract MLHO tables from Phase 2.1 tables.
 
```{r}
#uncomment the following line to execute
 # labeldt <- PatientSummary %>% dplyr::select(patient_num,label=severe) %>% mutate(label=factor(label))
 
 #dbmart <- PatientObservations %>% 
  #dplyr::filter(days_since_admission >= 0) %>%
  #dplyr::filter(days_since_admission %in% severcode$concept_cd) %>%
  #dplyr::select(patient_num,start_date=days_since_admission,phenx=concept_code)
  
#dems<-as.data.frame(model.matrix(~ . + 0, data=PatientSummary[,c(1,10,11,12)], contrasts.arg = lapply(PatientSummary[,c(10,11,12)],
#contrasts, contrasts=FALSE)))

#dems$sexMale <- NULL

```

## Step 2: Implementation

Now we update the list of unique patients.
```{r}
#uncomment the following line to execute
#uniqpats <- c(as.character(unique(dbmart$patient_num)))
```

step 4- we create a wide table for analysis using the `wideout` function
```{r}
#uncomment the following line to execute
#dbmart.wide <- wideout(data=dbmart,patients=uniqpats)
```

step 5- the ML part!
the following function will do an iterative run of the GLM boost and stores the results in the `results` directory.
```{r}
#uncomment the following line to execute
# learn(data= dbmart.wide,
#       demographics=dems,
#       augment=FALSE,
#       tst.size=0.2,
#       save.model=FALSE,
#       note="mlho_terst_run",
#       aoi="Severity",
#       multicore=TRUE)
```


### Next, we will visualize the results.
